---
title: "NLPReviewClassification"
author: "SantaBorel"
date: '2018-07-09'
output:
  word_document: default
  pdf_document: default
---

# Restaurant Comment Review Classsification using NLP

## Introduction 
Yelp is a company that hosts a community where users review local businesses that they have visited.
Users can check in and leave reviews for businesses that they have visited by giving it a star rating and
writing a description. These reviews can then be voted on my other users. Yelp has made their data
available to the public.  
I am interested in evaluating these reviews using Natural language processing (NLP) and other
techniques to be able to predict the star rating of restaurant reviews. Even though on the Yelp platform
users are required to report a star rating, on other platforms it is not required (e.g. blog posts, Tweets,
Instagram comments). For businesses to understand what clients think of them, it is important to
evaluate multiple platforms and this will include ones where star ratings are not provided and the
sentiment of the review will need to be interpreted. If a good algorithm is developed it could be used to
sort through reviews/comments to quickly evaluate how clients are enjoying their business.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(jsonlite)
library(caret)
library(stopwords)
library(text2vec)
library(glmnet)
library(ggplot2)
```

### Loading the business.json file from the Yelp data
This file provides information about the businesses included in Yelp.

```{r loading business data}
business <- stream_in(file("dataset/business.json"), verbose = FALSE)

# Filter on businesses that are in the Canadian province of Ontario
businessON <- business[which(business$state=="ON"),];

# Remove full business file to save memory
rm(business)

```

### Explore the businessON data files
This analysis is focused on the restaurant ratings. Then we need to determine which fields and values can be used to filter the data to just restaurant establishments.
It is important to filter the data to just the ones that we are interested in. NLP models are very contex dependement since people may speak differently or leave different types of phrases for different categories of businesses.

```{r get data for restaurants, echo=FALSE}

str(businessON)

# First explore the types of categories that could be associated with a restaurant/food serving business
categories <- data.frame(table(unlist(businessON$categories)))
# Order by the count
categ.ordered <- categories[order(categories$Freq, decreasing = TRUE),]

head(categ.ordered, 20)
# A lot of the potentially food related categories are ambiguous so only the businesses containing
# the 'Restaurant' tag for the Categories field will be included

restON <- businessON[mapply('%in%',"Restaurants",businessON$categories),];

# Remove businessON data to save memory
rm(businessON)

```

### Loading the review.json file from the Yelp data
The next step is to load the data that contain the user text reviews and other review parameters.
The full file is too large to be handled in memory so the loading method filters the data to save the reviews that have the business_id that is included in restON (restaurants from Ontario)

```{r load reviews}

#Pull in reviews where businesses are in ON and a Restaurant
con_in <- file("dataset/review.json")
con_out <- file(tmp <- tempfile(), open = "wb")
stream_in(con_in, handler = function(df){
      df <- df[which(df$business_id %in% restON$business_id),]
      stream_out(df, con_out, pagesize = 1000, verbose = FALSE)
}, pagesize = 5000)
close(con_out)
reviews <- stream_in(file(tmp))
nrow(reviews)
unlink(tmp)

str(reviews)

```

## Evaluate the reviews rating distribution 
Look at the distribution for the number of stars given to businesses. The initial analysis will focus on being able to classify a review as good or bad (e.g. a binary score). This is defined as below or above median

```{r evaluate star distribution}

## Get distribution for number of reviews
table(as.numeric(reviews$stars))
hist(as.numeric(reviews$stars))

# there are no null values here - all reviews have a star rating
sum(is.na(as.numeric(reviews$stars)))

# For the initial analysis the scoring will be simplified to a binary value - below or above the median

median(as.numeric(reviews$stars))
reviews$starsbinary <- NA;
reviews$starsbinary[which(as.numeric(reviews$stars) <= 3)] <- 0;
reviews$starsbinary[which(as.numeric(reviews$stars) > 3)] <- 1;

head(reviews$stars,10)
head(reviews$starsbinary,10)

```

## Creating a Model
Now that the data is prepared the model can start to be prepared. 

### Prepare training and test sets

```{r train test sets}
# Seed is set for reproducibility
set.seed(100)

# From the caret package the createDataPartition function is used to created indices for 
# test (40%) and training (60%) data. The data is sampled based on the created binary rating
# field (starsbinary)
trainIndex = createDataPartition(reviews$starsbinary, p=0.6, list=FALSE)
train = reviews[trainIndex,]
test = reviews[-trainIndex,]

```

### Prepare text data
First the text data needs to be prepared to be ingested by an algorithm.
The text reviews are tokenized and a document term matrix is created.

```{r tokenize and dtm}

# Used to create iterators for further use
# The text data will be tokenized to words and all text will be made lowercase
# Progress bar is set to false since it doesn't display well in rmd
t1 <- itoken(train$text,
             preprocessor = tolower,
             tokenizer = word_tokenizer,
             ids = train$review_id,
             progressbar = FALSE)

# Maybe consider taking out numbers.

# The vocabulary matrix is created with stopwords being removed
vocab <- create_vocabulary(t1, stopwords = stopwords::stopwords("en"))
head(vocab, 25)
```

There are a lot of words that are rarely used in reviews and that may be misspelled or in a non-english language. Since NLP will preform best when only one language is used, these need to be filtered out. Less commonly used words are also removed since they will just add noise to the algorithm.

```{r}
vocab_pruned <- prune_vocabulary(vocab, 
                          term_count_min = 10, 
                          doc_proportion_max = 0.5,
                          doc_proportion_min = 0.001)

vectorizer = vocab_vectorizer(vocab_pruned)

# A document term matrix is created
dtm_train = create_dtm(t1, vectorizer) 
dim(dtm_train) # rows are the reviews and columns are the terms

# for example
# dtm_train[1,which(dtm_train[1,] != 0)]

```

### Create Linear Model

The glmnet package will be used to create a linear model. This package uses extrememly efficient procedures for fitting data (by fitting the entrie lasso or elastic-net reguarization path). It uses a penalized maximum likelihood. It is a fast algorithm that can exploit sparcity in the input matrix. This works well with the document term matrix since it is spacely populated.  
The cv.glmnet function used here, uses cross-validation(cv). The cv used here was 4-fold.

```{r glmnet train}

time_initial = Sys.time()

glmnet_classifier = cv.glmnet(x = dtm_train, y = train$starsbinary,
                              # Binomial model is used for the fit
                              family = 'binomial',
                              # L1 penalty, alpha 1 is the lasso penalty
                              alpha = 1,
                              # interested in the area under ROC curve
                              type.measure = "auc",
                              # n-fold cross-validation
                              nfolds = 4,
                              # Convergence threshold. 
                              # Higher value is less accurate, but has faster training
                              thresh = 1e-3,
                              # Maximum number of itterations.
                              # Lower value for faster training
                              maxit = 1e3)

# Print the time it took
print(difftime(Sys.time(), time_initial, units = 'min'))

# Plot the AUC value
plot(glmnet_classifier)

# The maximum AUC value
print(paste("max AUC =", round(max(glmnet_classifier$cvm), 4)))

```

### Testing the Model

```{r glmnet test}

# The iterators will be created for the test data
it_test <- itoken(test$text,
             preprocessor = tolower,
             tokenizer = word_tokenizer,
             ids = test$review_id,
             progressbar = FALSE)

# The DTM is created using the vectorizer from the training data evaluation
dtm_test = create_dtm(it_test, vectorizer)

# Again the DTM has rows equal to the number of reviews in the test data and 
# the columns are all the terms
dim(dtm_test)

# Get the predictions for the test data using the glmnet_classifier
preds = predict.cv.glmnet(glmnet_classifier, dtm_test, type = 'response')[,1]
auctest = glmnet:::auc(test$starsbinary, preds)

# The maximum AUC value for the test data
print(paste("max AUC test =", round(max(auctest), 4)))

CM <- confusionMatrix(as.factor(round(preds,0)), as.factor(test$starsbinary))
print(paste("Accuracy =", round(CM$overall[1], 3)))

resultsForPlotting <- data.frame('ID' = test$review_id,
                                'Pred' = preds,
                                'Actual' = test$starsbinary,
                                'ActualStars' = test$stars)
h <- ggplot(data = resultsForPlotting, aes(Pred)) + 
      geom_histogram(aes(fill = as.factor(Actual)), bins = 20)
h

h2 <- ggplot(data = resultsForPlotting, aes(x=Pred, fill = as.factor(ActualStars))) + 
      geom_density(alpha = 0.1)
h2

```

The values for the test data is close to the training data which indicates that the model did not overfit.
The results give a high AUC=0.924

### Normalization
Now the same model will be used but with normalized data. L1 Normalization is applied where the transformation is done in a way so that the sum of each row (review) will be equal to 1. This 

```{r normalize and remodel}

dtm_train_norm = normalize(dtm_train, "l1")

glmnet_classifier_norm = cv.glmnet(x = dtm_train_norm, y = train$starsbinary,
                              family = 'binomial',
                              alpha = 1,
                              type.measure = "auc",
                              nfolds = 4,
                              thresh = 1e-3,
                              maxit = 1e3)

# The maximum AUC value
print(paste("max AUC =", round(max(glmnet_classifier_norm$cvm), 4)))

dtm_test_norm = normalize(dtm_test, "l1")

# Get the predictions for the test data using the glmnet_classifier
preds = predict.cv.glmnet(glmnet_classifier_norm, dtm_test_norm, type = 'response')[,1]
auctest = glmnet:::auc(test$starsbinary, preds)

# The maximum AUC value for the test data
print(paste("max AUC test =", round(max(auctest), 4)))

CM <- confusionMatrix(as.factor(round(preds,0)), as.factor(test$starsbinary))
print(paste("Accuracy =", round(CM$overall[1], 3)))

resultsForPlotting <- data.frame('ID' = test$review_id,
                                'Pred' = preds,
                                'Actual' = test$starsbinary,
                                'ActualStars' = test$stars)
h <- ggplot(data = resultsForPlotting, aes(Pred)) + 
      geom_histogram(aes(fill = as.factor(Actual)), bins = 20)
h

h2 <- ggplot(data = resultsForPlotting, aes(x=Pred, fill = as.factor(ActualStars))) + 
      geom_density(alpha = 0.1)
h2

```

Normalization provided slightly better results at AUC = 0.930 for both test and train data.

### TF_IDF
Now TF-IDF will be used to transform the training and test sets

```{r TF-IDF}

# define a TF-IDF model object
tfidf_model = TfIdf$new()

# fit the model to the training data
dtm_train_tfidf = fit_transform(dtm_train, tfidf_model)

# apply model to train data to prepare tf-idf version
dtm_test_tfidf = transform(dtm_test, tfidf_model)

time_initial = Sys.time()
glmnet_classifier_tfidf = cv.glmnet(x = dtm_train_tfidf, y = train$starsbinary,
                              family = 'binomial',
                              alpha = 1,
                              type.measure = "auc",
                              nfolds = 4,
                              thresh = 1e-3,
                              maxit = 1e3)
print(difftime(Sys.time(), time_initial, units = 'min'))

# The maximum AUC value
print(paste("max AUC =", round(max(glmnet_classifier_tfidf$cvm), 4)))

# Get the predictions for the test data using the glmnet_classifier
preds = predict.cv.glmnet(glmnet_classifier_tfidf, dtm_test_tfidf, type = 'response')[,1]
auctest = glmnet:::auc(test$starsbinary, preds)

# The maximum AUC value for the test data
print(paste("max AUC test =", round(max(auctest), 4)))

CM <- confusionMatrix(as.factor(round(preds,0)), as.factor(test$starsbinary))
print(paste("Accuracy =", round(CM$overall[1], 3)))

resultsForPlotting <- data.frame('ID' = test$review_id,
                                'Pred' = preds,
                                'Actual' = test$starsbinary,
                                'ActualStars' = test$stars)
h <- ggplot(data = resultsForPlotting, aes(Pred)) + 
      geom_histogram(aes(fill = as.factor(Actual)), bins = 20)
h

h2 <- ggplot(data = resultsForPlotting, aes(x=Pred, fill = as.factor(ActualStars))) + 
      geom_density(alpha = 0.1)
h2

```

The results with glm using tf-idf preprosessed data produced similar reults to the Normalized data with AUC=0.930

```{r plot some results}

resultsForPlotting <- data.frame('ID' = test$review_id,
                                'Pred' = preds,
                                'Actual' = test$starsbinary,
                                'ActualStars' = test$stars)
h <- ggplot(data = resultsForPlotting, aes(Pred)) + 
      geom_histogram(aes(fill = as.factor(Actual)), bins = 20)
h

h2 <- ggplot(data = resultsForPlotting, aes(x=Pred, fill = as.factor(ActualStars))) + 
      geom_density(alpha = 0.1)
h2

h3 <- ggplot(data = resultsForPlotting, aes(x=Pred, fill = Actual)) + 
      geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity', bins = 20)
h3

```

### GloVe Word Embeddings

The same test set vectorizer will be used as above. Now instead of creating a document term matrix I will construct a term-co-occurence matrix (TCM). 
This did not work for now.

```{r word embedding}

## The same test set vectorizer will be used as above.

# tcm <- create_tcm(t1, vectorizer, skip_grams_window = 5L)
# 
# time_initial = Sys.time()
# 
# # The TCM matrix will be facorized using hte GloVE algorithm
# glove = GlobalVectors$new(word_vectors_size = 50, vocabulary = vocab_pruned, x_max = 10)
# glove$fit_transform(tcm, n_iter = 20)
# 
# # Print the time it took
# print(difftime(Sys.time(), time_initial, units = 'min'))
# 
# word_vectors <- glove$components
# 
# wordtest <- word_vectors["good", ] - 
#   word_vectors["service", ] + 
#   word_vectors["restaurant", ]
# cos_sim = sim2(x = word_vectors, y = wordtest, method = "cosine", norm = "l2")
# head(sort(cos_sim[,1], decreasing = TRUE), 5)

```

### Naive Bayes Classifier

```{r Naive Bayes}

library(klaR)
time_initial = Sys.time()
nb_classifier = klaR::NaiveBayes(as.matrix(dtm_train), as.factor(train$starsbinary), fL = 1)
print(difftime(Sys.time(), time_initial, units = 'min'))

pred_nb <- predict(nb_classifier, newdata=as.matrix(dtm_train))

``` 



```{r svm}

library(e1071)
time_initial = Sys.time()
svm_classifier = svm(as.matrix(dtm_train)[1:10000,], train$starsbinary[1:10000])
print(difftime(Sys.time(), time_initial, units = 'min'))

pred_nb <- predict(nb_classifier, newdata=as.matrix(dtm_train))

``` 